---
title: "Course Project"
author: "Aasish Agarwal"
date: "Sunday, December 07, 2014"
output: html_document
runtime: shiny
---

## Objective

The purpose of this project is to demonstrate your ability to collect, work with, and clean a data set. The goal is to prepare tidy data that can be used for later analysis. You will be graded by your peers on a series of yes/no questions related to the project. You will be required to submit: 

1) a tidy data set as described below, 
2) a link to a Github repository with your script for performing the analysis, and 
3) a code book that describes the variables, the data, and any transformations or work that you performed to clean up the data called CodeBook.md. You should also include a README.md in the repo with your scripts. This repo explains how all of the scripts work and how they are connected.  

One of the most exciting areas in all of data science right now is wearable computing - see for example this article . Companies like Fitbit, Nike, and Jawbone Up are racing to develop the most advanced algorithms to attract new users. The data linked to from the course website represent data collected from the accelerometers from the Samsung Galaxy S smartphone. A full description is available at the site where the data was obtained: 

http://archive.ics.uci.edu/ml/datasets/Human+Activity+Recognition+Using+Smartphones 

Here are the data for the project: 

https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip 

 You should create one R script called run_analysis.R that does the following. 

1. Merges the training and the test sets to create one data set.
2. Extracts only the measurements on the mean and standard deviation for each measurement. 
3. Uses descriptive activity names to name the activities in the data set
4. Appropriately labels the data set with descriptive variable names. 
5. From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each subject.

Good luck!


### Downloading Data
```{r, echo=FALSE}
setwd("D:\\dev\\datascience\\gncdata")

url <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
  if (! file.exists("data")){
    dir.create("data")
  } 
  projectDatafile <- "data/Dataset.zip"

  cat ("Operating in folder: " , getwd())
  cat("Downloadoing data set if not already exists: " , projectDatafile)
  if (! file.exists(projectDatafile)){
      cat("Staring data download")
      download.file(url , projectDatafile, method = "curl")
  } else {
    cat("File already downloaded")
  } 
```
### Uncompressing Data
```{r, echo=TRUE}
  library(R.utils)
  created_dir <- "UCI HAR Dataset"
  if ( ! file.exists(created_dir)){
    cat("Uncompressing file: ", projectDatafile)
    unzip(projectDatafile)
  } else {
    cat("File already uncompressed")
  } 
  cat("Done")

```
### Loading Data Sets
We have a total of 8 data sets to load

- 'features.txt': List of all features. These are the names of 561 different indicators available in test & traing data sets.
- 'activity_labels.txt': Links the class labels with their activity name.
- 'train/subject_train.txt': Each row identifies the subject who performed the activity for each window sample. Its range is from 1 to 30. 
- 'train/X_train.txt': Training set.
- 'train/y_train.txt': Training Activity Id.
- 'train/subject_test.txt': Each row identifies the subject who performed the activity for each window sample. Its range is from 1 to 30. 
- 'test/X_test.txt': Test set.
- 'test/y_test.txt': Test Activity Id.

```{r, echo=TRUE}
library(dplyr)
featurenames <- read.table(file.path(created_dir,"features.txt")) %>% tbl_df()
activitylabels <- read.table(file.path(created_dir,"activity_labels.txt")) %>% tbl_df()

training_subjects <- read.table(file.path(created_dir,"train", "subject_train.txt")) %>% tbl_df()

training_activities <- read.table(file.path(created_dir,"train", "y_train.txt")) %>% tbl_df()

training_data <- read.table(file.path(created_dir,"train", "X_train.txt")) %>% tbl_df()

test_subjects <- read.table(file.path(created_dir,"test", "subject_test.txt")) %>% tbl_df()

test_activities <- read.table(file.path(created_dir,"test", "y_test.txt")) %>% tbl_df()

test_data <- read.table(file.path(created_dir,"test", "X_test.txt")) %>% tbl_df()

cat("featurenames: ", nrow(featurenames),
    ", activitylabels: ", nrow(activitylabels),
    ", training_subjects: ", nrow(training_subjects),
    ", training_activities: ", nrow(training_activities),
    ", training_data: ", nrow(training_data),
    ", test_subjects: ", nrow(test_subjects),
    ", test_activities: ", nrow(test_activities),
    ", test_data: ", nrow(test_data)
    )

```

#### Step1 - Merging Data Sets

We will merge 

* test & training subjects
* test & training activity id
* test & training data sets

```{r, echo=TRUE}
activitives <- rbind(training_activities, test_activities)
subjects <-  rbind(training_subjects, test_subjects)
dataset <- rbind(training_data, test_data)

cat("activitives: ", nrow(activitives),
    ", subjects: ", nrow(subjects),
    ", dataset: ", nrow(dataset)
    )
```

#### Step2 - Extracting mean & standard deviation

Add names to merged data set
Update data set to include only mean & standard deviation

```{r, echo=TRUE}
names(dataset) <- featurenames$V2
colnames <- grepl("mean\\(\\)|std\\(\\)", names(dataset))
dataset <- dataset[colnames]

```

#### Step3 - Use descriptive activity names

Update Activity to change Id with descriptive value
Join Activity, Data Set

```{r, echo=TRUE}
names(activitives) <- c("activityid")
names(activitylabels) <- c("activityid","activity")
names(subjects) <- c("subject")
activitives <- merge(activitives, activitylabels, by="activityid", all.x=TRUE) %>% tbl_df()

dataset <- cbind(activitives,dataset)
dataset <- cbind(subjects,dataset) %>% 
  tbl_df() %>% 
  select(-activityid)
dataset
```

#### Step4 - Label data set with descriptive variable names

Join Activity, Subject, Data Set

#### Step5 - Create a second, independent data set with average of each variable for each activity & each subject














