---
title: "Getting & Cleaning Data - Quiz 1"
author: "Aashsih"
date: "Friday, November 14, 2014"
output: html_document
runtime: shiny
---

## Question 1
The American Community Survey distributes downloadable data about United States communities. Download the 2006 microdata survey about housing for the state of Idaho using download.file() from here: 

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv 

and load the data into R. The code book, describing the variable names is here: 

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FPUMSDataDict06.pdf 

How many properties are worth $1,000,000 or more?

### Answer
```{r, echo=FALSE}
  url <- "http://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv"
  if (! file.exists("data")){
    dir.create("data")
  } 
  datafile <- "data/housingdata.csv"

if (! file.exists(datafile)){
    download.file(url , datafile)
  } 


  codebookurl <- "http://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FPUMSDataDict06.pdf"
  codebookfile <- "data/housingdata-codebook.pdf"
  if (! file.exists(codebookfile)){
      download.file(codebookurl , codebookfile)
    } 

  housingdata <- read.csv(datafile)
  state_code <- 16 
  prop_value_coe <- 24
  high_value_properties <- subset(housingdata, housingdata$ST == state_code & housingdata$VAL == prop_value_coe)
sprintf("There are %s properties which are worth $1,000,000 or more in Idoho ", nrow(high_value_properties))
```

## Question 2
Use the data you loaded from Question 1. Consider the variable FES in the code book. Which of the "tidy data" principles does this variable violate?

### Answer
Data contains two variables "*Family type and employment status*"
This violates "Tidy data has one variable per column." tidy data prunciple

## Question 3
Download the Excel spreadsheet on Natural Gas Aquisition Program here: 

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FDATA.gov_NGAP.xlsx 

Read rows 18-23 and columns 7-15 into R and assign the result to a variable called:
 dat 
What is the value of:
 sum(dat$Zip*dat$Ext,na.rm=T) 

### Answer
```{r, echo=FALSE}
  url <- "http://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FDATA.gov_NGAP.xlsx"
  if (! file.exists("data")){
    dir.create("data")
  } 
  datafile <- "data/gov_NGAP.xlsx"

if (! file.exists(datafile)){
    download.file(url , datafile)
  } 

library(xlsx)
colIndex <- 7:15
rowIndex <- 18:23
dat  <- read.xlsx(file = "data/gov_NGAP.xlsx", sheetIndex = 1, colIndex = colIndex, rowIndex = rowIndex)
sum(dat$Zip*dat$Ext,na.rm=T) 

```
## Question 4
Read the XML data on Baltimore restaurants from here: 

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Frestaurants.xml 

How many restaurants have zipcode 21231?

### Answer

```{r, echo=FALSE}
library(XML)
  url <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Frestaurants.xml"
  if (! file.exists("data")){
    dir.create("data")
  } 
  datafile <- "data/restaurants.xml"

if (! file.exists(datafile)){
    download.file(url,destfile = datafile , method = "curl")
  } 

doc <- xmlTreeParse("data/restaurants.xml",useInternalNodes = TRUE)
rootNode <- xmlRoot(doc)
allzipcodes <- xpathSApply(rootNode,"//zipcode",xmlValue)
required_zip <- allzipcodes == "21231"
length(allzipcodes[required_zip])

```
## Question 5
The American Community Survey distributes downloadable data about United States communities. Download the 2006 microdata survey about housing for the state of Idaho using download.file() from here: 

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06pid.csv 

using the fread() command load the data into an R object
 DT 
Which of the following is the fastest way to calculate the average value of the variable
pwgtp15 
broken down by sex using the data.table package?
### Answer

```{r, echo=FALSE}
library(XML)
  url <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06pid.csv"
  if (! file.exists("data")){
    dir.create("data")
  } 
  datafile <- "data/uscommunities.csv"

if (! file.exists(datafile)){
    download.file(url,destfile = datafile , method = "curl")
  } 
  DT <- read.csv(datafile)

tapply(DT$pwgtp15,DT$SEX,mean)
sapply(split(DT$pwgtp15,DT$SEX),mean)
mean(DT$pwgtp15,by=DT$SEX)
mean(DT[DT$SEX==1,]$pwgtp15); mean(DT[DT$SEX==2,]$pwgtp15)

```


## About The Document

This R Markdown document is made interactive using Shiny. Unlike the more traditional workflow of creating static reports, you can now create documents that allow your readers to change the assumptions underlying your analysis and see the results immediately. 

To learn more, see [Interactive Documents](http://rmarkdown.rstudio.com/authoring_shiny.html).

## Inputs and Outputs

You can embed Shiny inputs and outputs in your document. Outputs are automatically updated whenever inputs change.  This demonstrates how a standard R plot can be made interactive by wrapping it in the Shiny `renderPlot` function. The `selectInput` and `sliderInput` functions create the input widgets used to drive the plot.

```{r, echo=FALSE}
inputPanel(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  
  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")
  
  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```

## Embedded Application

It's also possible to embed an entire Shiny application within an R Markdown document using the `shinyAppDir` function. This example embeds a Shiny application located in another directory:

```{r, echo=FALSE}
shinyAppDir(
  system.file("examples/06_tabsets", package="shiny"),
  options=list(
    width="100%", height=550
  )
)
```

Note the use of the `height` parameter to determine how much vertical space the embedded application should occupy.

You can also use the `shinyApp` function to define an application inline rather then in an external directory.

In all of R code chunks above the `echo = FALSE` attribute is used. This is to prevent the R code within the chunk from rendering in the document alongside the Shiny components.



